---
title: "Assignment 1"
author: 
  -   "Pablo Vidal Fernández"
  -   "Cripto Ignacio Díez Ruiz"
  -   "Jialian Zhou He"
  -   "Carlos Roldán Piñero"
  
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, 
                      message = F)
```

```{r, echo = F, results = "hide"}
if(!require(fpp2)){
  install.packages("fpp2")
}

if(!require(fpp3)){
  install.packages("fpp3")
}

if(!require(ggplot2)){
  install.packages("ggplot2")
}

if(!require(ggpubr)){
  install.packages("ggpubr")
}

library(fpp2)
library(fpp3)
library(ggplot2)
library(ggpubr)
```


## Exercise 1 

Use the appropiate graphics functions, explore features from the following time series: **bicoal**, **bricksq**, **hsales**, **ibmclose**, **Internet**, **writing**.

-   Can you spot any seasonality, cyclicity and trend?

-   What do you learn about the time series?

-   Justify the choice of the graphic function. 

### Bicoal TS

The first thing that we have to check is the frequency of the time series:

```{r}
frequency(bicoal)
```
It is annual data, therefore it cannot be seasonal.

Plotting the series and the first differences:
```{r, fig.cap="Bicoal"}
p1 <- autoplot(bicoal, main = "Bicoal TS", xlab = "Year", ylab = "Bicoal")
p2 <- autoplot(diff(bicoal), main = "Bicoal TS", xlab = "Year", ylab = "Lagged bicoal")
ggarrange(p1, p2, ncol = 2)
```

We can see that there is no overall trend, and there seems to be a cyclic behaviour every 10 years. The plot for the differences resembles white noise. At this point, one could suspect stationarity.

```{r}
ggAcf(bicoal, lag.max = 45)
```
We can observe a sinusoidal pattern, and it goes towards zero. It seems that this time series is stationary, with a cyclic component.


### Bricksq TS

```{r}
frequency(bricksq)
```
We have quarterly data. Thus, seasonality can exist.

```{r}
par(mfrow=c(1,2))
p3 <- autoplot(bricksq, main = "Bricksq TS", xlab = "Year", ylab = "Bricksq")
p4 <- autoplot(diff(bricksq, lag = 1), main = "Bricksq TS", xlab = "Year", ylab = "Lagged Bricksq")
ggarrange(p3, p4, ncol = 2)

```

Looking at the plot, we can see that there might be a seasonal component and a trend. The series doesn't appear to be stationary.

```{r}
ggAcf(bricksq)
```

As the autocorrelations for the first lags are large and they slowly decrease, we can say that there is a trend. Moreover, we can see that each 4 lags the autocorrelation is higher, a sign of seasonality.

```{r}
p5 <- ggseasonplot(bricksq)
p5
```
There are a few outliers in the Q4 for some years, and all years seem to present a very similar pattern: Q2 and Q3 are higher than Q1 and Q4.

We can appreciate that pattern more clearly in the subseries plot:

```{r}
p6 <- ggsubseriesplot(bricksq)
p6
```

### Hsales TS ###

```{r}
frequency(hsales)
```
We have monthly data, thus a seasonal component might exist. 

```{r}
p7 <- autoplot(hsales, main = "Hsales TS")
p8 <- autoplot(diff(hsales, lag = 1), main = "Bricksq TS", xlab = "Year", ylab = "Lagged Hsales")

ggarrange(p7, p8, ncol = 2)
```

At a first glance, we can see a cyclical component in the series. There doesn't appear to be any trend. Moreover, a seasonal component can be appreciated within each year.

```{r}
p9 <- ggseasonplot(hsales)
p9
```

We can see that march and april tend to be the months with higher volume, whereas the last months of the year come with a decrease. In August, there seems to be a recuperation with respect to July.

```{r}
p10 <- ggAcf(hsales)
p10
```


There are peaks at 12 and 24 in the correlogram, confirming seasonality. There is no evidence for a trend.

### Ibmclose TS

```{r}
frequency(ibmclose)
```

The data is daily.

```{r}
autoplot(ibmclose)
```

The data is daily, so there can't be a seasonal nor cyclic component. It doesn't look like there is a trend either. 

```{r}
ggAcf(ibmclose, lag.max = 50)
```

The series is definitely not stationary. Correlations are very high and don't go rapidly towards zero.

### Internet TS

```{r}
frequency(internet)
autoplot(internet)
```

Using *help*, we can see that the dataset contains the number of user logged on to an internet server each minute over a 100-minute period. Consequently, there won't be a cyclic or seasonal component.

It doesn't look like there is a trend. The series goes up, then goes down and finally it starts rising again. 

```{r}
ggAcf(internet, lag.max = 90)
```
We can see oscillations between positive and negative autocorrelations, consistent with the ups and downs in the series. It doesn't go rapidly towards zero, so the data is not stationary.


### Writing TS

```{r}
frequency(writing)
```
We have monthly data. 

```{r}
autoplot(writing)
```

We can see a positive linear trend, a strong seasonality and there doesn't seem to be a cyclic component.

```{r}
ggseasonplot(writing)
```
In this plot, we can appreaciate that each year, values keep getting higher, so there is a positive trend. Furthermore, there is always a big drop in August, possibly related to the vacation period of workers.

```{r}
ggsubseriesplot(writing)
```

We can see the same information in the subseries plot.

```{r}
autoplot(diff(writing))
```
We can see in the first differences series that variability also grows as time passes.

```{r}
ggAcf(writing)
```
Peaks at 12 and 24 confirm the seasonality.

## Exercise 5

The data below represent the monthly sales (in thousands) of product A for a plastics manufacturer for years 1 through 5 (data set plastics).

(a) Plot the time series of sales of product A. Can you identify any key feature? Explain what you
see.

```{r}
autoplot(plastics)
```

We can see a strong seasonality component, with an incremental production up to the summer, and then a decrease; and also a positive and linear trend. 

(b) Use an STL decomposition to calculate the trend-cycle and seasonal indices. (Experiment with
having fixed or changing seasonality). Do you see any difference?

```{r}

```

(c) Please, discuss whether the results support the graphical interpretation form part (a).

```{r}

```

(d) Compute and plot the seasonally adjusted data.

```{r}

```

(e) Use a random walk to produce forecasts of the seasonally adjusted data.

```{r}

```

(f) Re-seasonilize the results to give forecasts on the original scale.

```{r}

```

